```{r}
if (!require("readxl")) install.packages("readxl", repos='http://cran.us.r-project.org')
if (!require("writexl")) install.packages("writexl", repos='http://cran.us.r-project.org')
if (!require("dplyr")) install.packages("readxl", repos='http://cran.us.r-project.org')
# install.packages("ggplot2", repos='http://cran.us.r-project.org')

remotes::install_version('ggplot2', version='3.5.1')
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.19")
BiocManager::install("ggtree")
BiocManager::install("BiocFileCache")
BiocManager::install("clusterProfiler") 
BiocManager::install("org.Hs.eg.db")
BiocManager::install("AnnotationDbi")
BiocManager::install("biomaRt")

```
# load libraries
```{r}
library(readxl)
library(writexl)
library(dplyr)
library(ggplot2)
library(RColorBrewer)

library(BiocFileCache)
library(clusterProfiler)
library(org.Hs.eg.db)
library(AnnotationDbi)
library(biomaRt)

```

# check the data
```{r}
# Read the Excel file
data <- read_excel("Gene_level_predictions.xlsx")

# Check for missing values
print(colSums(is.na(data)))

# View the structure of the data
str(data)

# Remove rows with missing values
data <- na.omit(data)

# check how many data left
str(data)
```

# classify the tags

```{r}
# Create a new column to store the classification result
data$Group <- NA

# Classification logic: classify based on N score, C score, and Internal score
# Check if column names are correct and handle missing values
if (!("Internal score" %in% colnames(data)) || !("N score" %in% colnames(data)) || !("C score" %in% colnames(data))) {
  stop("Column names do not match or are missing.")
}

for (i in 1:nrow(data)) {
  if (!is.na(data$`Internal score`[i]) && data$`Internal score`[i] >= data$`N score`[i] && data$`Internal score`[i] >= data$`C score`[i]) {
    data$Group[i] <- "Internal"
  } else if (!is.na(data$`N score`[i]) && data$`N score`[i] >= data$`C score`[i]) {
    data$Group[i] <- "N terminus"
  } else if (!is.na(data$`C score`[i])) {
    data$Group[i] <- "C terminus"
  }
}

```

# filter out datas depending on number of sequences

```{r}
# Set a threshold, e.g., filter out rows with Number of sequences less than 10
# threshold <- 10
# data <- data[data$`Number of sequences` >= threshold, ]
```

# output to grouped data to file
```{r}
# Sort data by Group (N, C, Internal)
data_sorted <- data %>%
  mutate(Group = factor(Group, levels = c("N terminus", "C terminus", "Internal"))) %>%
  arrange(Group)

# Save the sorted data to an Excel file
write_xlsx(data_sorted, "Gene_level_grouped.xlsx")

# write three grouped into three files
list_of_groups <- unique(data_sorted$Group)

for (group in list_of_groups) {
  group_data <- filter(data_sorted, Group == group)
  group_name_no_space <- gsub(" ","",group)
  file_name <- paste("Gene_level_", group_name_no_space, ".xlsx", sep = "")
  write_xlsx(group_data,file_name)
}
```

# calculate percentage of the tagging sites(N/C/Internal)
```{r}
# calculate percentage
group_counts <- data %>%
  group_by(Group) %>%
  summarise(Count = n()) %>%
  mutate(Percentage = round((Count / sum(Count) * 100),2))

print(group_counts)

# print(colnames(data))
```
# Descriptive statistical analysis
## Plotting Percentage Stacked Bars: percentage vs group
```{r}

percentage_plot <- ggplot(group_counts, aes(x = Group, y = Percentage, fill = Group)) +
  geom_col() +
  geom_label(aes(label = paste(Count, " (", Percentage, "%)", sep = "")),# add percentage label
            position = position_stack(vjust = 0.5), # center the label
            fill = "white", # background color
            color = "black", # font color
            size = 3.5) + # font size
  labs(title = "Percentage of Numbers by Group",
       x = "Group",
       y = "Number of Genes") +
  theme_minimal() +
  scale_fill_brewer(palette = "Dark2")

percentage_plot

```
## plot binned histogram -- Distribution of Number of Sequences (Internal)
```{r}
data_internal <- read_excel("Gene_level_Internal.xlsx")
data_N <- read_excel("Gene_level_Nterminus.xlsx")
data_C <- read_excel("Gene_level_Cterminus.xlsx")

binned_his <- ggplot(data_internal, aes(x = `Number of sequences`)) +
  geom_histogram(binwidth = 500, fill = "#6A5ACD", color = "#6A5ACD") +
  labs(title = "Distribution of Number of Sequences (Internal)",
       x = "Number of Sequences",
       y = "Frequency") +
  theme_minimal() 

binned_his


p_intelnal <- ggplot(data_internal %>% mutate(SeqNlog = log(`Number of sequences`)), 
       aes(x=SeqNlog)) + 
        labs(x = "Log Number of Sequences (Internal)",
             y = "Frequency") +
 geom_histogram(aes(y=after_stat(density)), colour="black", fill="white", binwidth=0.3)+
 geom_density(alpha=.2, fill="#8B0000") +
theme_minimal()
 #theme_bw()

p_N <- ggplot(data_N %>% mutate(SeqNlog = log(`Number of sequences`)), 
       aes(x=SeqNlog)) + 
        labs(x = "Log Number of Sequences (N Terminal)",
             y = "Frequency") +
 geom_histogram(aes(y=after_stat(density)), colour="black", fill="white", binwidth=0.3)+
 geom_density(alpha=.2, fill="#5D478B") +
theme_minimal()
 #theme_bw()

p_C <- ggplot(data_C %>% mutate(SeqNlog = log(`Number of sequences`)), 
       aes(x=SeqNlog)) + 
        labs(x = "Log Number of Sequences (C Terminal)",
             y = "Frequency") +
 geom_histogram(aes(y=after_stat(density)), colour="black", fill="white", binwidth=0.3)+
 geom_density(alpha=.2, fill="#2E8B57") +
theme_minimal()
 #theme_bw()

p_intelnal
p_N
p_C

```
# Comparison between groups
## violin plot--distribution of M value
```{r}
violin_plot <- ggplot(data,aes(x = Group, y = `M value`, fill = Group)) +
  geom_violin(trim = FALSE)+
  geom_boxplot(width = 0.1, fill = "white", outlier.shape = NA)+
  scale_y_log10() +
  labs(title = "Distribution of M value by Group",
       x = "Group",
       y = "M value")+
  theme_minimal()+
  scale_fill_brewer(palette = "Dark2")
violin_plot
```
